---
title: "Handling the data import"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Import_ABF}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(nphys)
```

# Importing Data: Building the rda file. 

For this example we will be examining the `field.rda` data set in the demo directory. This data set

Most functions run in this are to be run on the `$ABF` data. It remains consistent, allowing functions to be applied to the nested list despite changes to the experimental paradigm


```{r}
# Import the abf files. 
Import_ABF(Slice ="field", dir = "demo/field")

```

The output file is named sA in your environment. This is a basic title that is meant to ease analysis portability.
sA contains a working dirctory `wd` where the data set is stored relative to the project directory. 
It lists the metadata associated with the experimental protocol, and it lists the files associated with the abf data.
The abf data is processed by a sorting function to name the data based upon the protocol being used.

```{r}
names(sA$ABF)
```

The experimental paradigm is centered around the conditioning stimulus or beginning of the drug wash in.

Drug wash in can be contaminated with baseline changes during wash, especially when observing field recoridngs.

Therefore, it is important to closely monitor changes to the response before assesing the effect. 

In the case of assesing synapic plasticity changes, the ideal experimental circumstance sees no changes to baseline before the induction protocol commences.

If changes are observed to the response during wash in, it is important to note whether these changes are associated with drug exposure or if they are products of the experimental process.


# Observing the data

Synaptic plasticity experiments are analyzed for the percent change of the rising phase of the post synaptic depolarization. 

```{r}



```













