---
title: "Handling the data import"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Import_ABF}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  fig.path = "figs"
)
library(nphys)
```

## Importing Data: Building the nnest.

In this vignette we will be importing the data associated with the `field` data set whose raw data is nested in the `exa` directory. This data set is an example of an experiment that assesed long-term depression at medial perforant path synpses. We are importing the data from files of the `.abf` format.  

```{r}
# Import the abf files. 
field = importABF(x ="field", dir = "exa/field", ret = TRUE, sv = FALSE)

```

If `ret = TRUE`, a list will return and you can name this as you like. The default return name is nnest. This basic ID is meant to serve as a way to forward data into analysis functions. The nnest contains a working dirctory `wd` that points to where data set is stored relative to the project directory. It also contains the `metadata` associated with the experimental protocol, and lists the files associated with the abf data.

```{r}
print(field$wd)
print(field$files)
```

The abf data is processed by a sorting function to name the data based upon the protocol being used.

```{r}
names(field$ABF)
```

The experimental paradigm is centered around the conditioning stimulus or beginning of the drug wash in.

```{r}
field$md$StimFile
```

Finally we are going to add some traces to our nnest for viewing simplicity using the built in functions `pullSweeps()` and `avgSweeps()`

```{r}
# We can utilize the dfs_ABF() function to extract the sampling interval 
sampleInt.ms <- unique(dfs_ABF(field$ABF, int = "samplingIntervalInSec", returnList = FALSE))*1000 # convert to ms
if(length(sampleInt.ms)>1)stop("Different sample frequencies")
# Ajustments can be made to the points before and after the stimArtifact identifier as to the range to be isolated.
adjNeg = -100
adjPos = 1400



field$traces <- list(

    ms = seq(sampleInt.ms, length.out = sum(abs(adjNeg), abs(adjPos),1), by =sampleInt.ms),

    Bl_Avg <- pullSweeps(field$ABF, pull = "PreC-Bl", adjNeg = adjNeg, adjPos = adjPos) %>%
        avgSweeps(),

    Decay_Avg <- pullSweeps(field$ABF, pull = "Decay", adjNeg = adjNeg, adjPos = adjPos) %>%
        avgSweeps(),

    Cond_Avg <- pullSweeps(field$ABF, pull = "Cond", adjNeg = adjNeg, adjPos = adjPos) %>%
        avgSweeps()
)
```


```{r}

xA = list(
  title = "Time (ms)"
)
yA = list(
  range = c(-1.2,0.1),
  title = "mV"
)



plot <-
  plotly::plot_ly(
    data = data.frame(field$traces),
    x = ~ ms,
    y = ~ Bl_Avg,
    type = "scatter",
    mode = "markers"
  ) %>%
  plotly::add_trace(p = plot,
                    x = ~ ms,
                    y = ~ Decay_Avg
                    ) %>% 
  plotly::add_trace(p = plot, 
                    x = ~ ms,
                    y = ~ Cond_Avg
                    ) %>% 
  

plot


```

