---
title: "nnest"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{nnest}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(nphys)
```

# The neurophysiology nest (nnest)

The purpose of nesting an experiment is to provide a concise roadmap that to 
supports active development and simplified interaction with the experiment and
the data that it is comprised of. The nnest is meant to be a cross paradigm simplification
that reduces each experiment to a collection of standardized elements that are 
critical for observing and analyzing the experiment and minimizing the impact of
variables such as acquisition software, file type & version, and metadata storage 
on the manipulation of the data. 


```{r}

```

## Standardized elements

Each nnest is expected to contain a an ID, a file directory, a metadata table, 
and a simplified roadmap for interacting with the raw data. Each nest can then also 
include analysis elements such as tables or matrices, and standardized figures for
observing the data.









